---
title: "STA457 Assignment 1"
author: "Yujia Zheng 1005329299"
date: "2021/1/21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
## (a)
```{r, fig.cap="Question 1a) Signal Plus Noise Model"}
set.seed(299) # get the same result each time
s1 <- rep(0,100) # s when t = 1,...,100
s2 <- 10*exp(-(101:200-100)/20)*cos(2*pi*101:200/4) # s when t = 101,...,200
s_a <- c(s1, s2) # combine s1 and s2
w <- rnorm(200, 0, 1) # generate Gaussian white noise
# plot the model
plot.ts(s_a+w, main = "Question 1a) Signal Plus Noise Model", col = "blue")
```
\newpage

## (b)
```{r, fig.cap="Question 1b) Signal Plus Noise Model"}
set.seed(299) # get the same result each time
s1 <- rep(0,100) # s when t = 1,...,100
s2 <- 10*exp(-(101:200-100)/200)*cos(2*pi*101:200/4) # s when t = 101,...,200
s_b <- c(s1, s2) # combine s1 and s2
w <- rnorm(200, 0, 1) # generate Gaussian white noise
# plot the model
plot.ts(s_b+w, main = "Question 1b) Signal Plus Noise Model", col = "blue")
```

## (c)
The appearance of series (a) is similar to the appearance of the explosion series in Fig. 1.7. The signal appears after a certain time in both plots, t=101 for plot(a) and t=1024 for the explosion series. They are both centered around 0, and the variance decreases with time.

The appearance of series (b) is similar to the appearance of the earthquake series in Fig. 1.7. The signal appears after a certain time in both plots, t=101 for plot(a) and t=1024 for the earthquake series. They are both centered around 0, and the variance decreases with time.

The first half of all four series are very similar. The variance of plot(b) & the earthquake series decreases slower than the variance of plot(a) & the explosion series.

```{r}
# plot the signal modulator
# signal modulator exp(-t/20)
plot.ts(exp(-(1:100)/20), main = "Signal Modulator exp(-t/20)", xlab = "t")
# signal modulator exp(-t/200)
plot.ts(exp(-(1:100)/200), main = "Signal Modulator exp(-t/200)", xlab = "t")
```

The signal modulator exp(-t/20) is more curved, it gradually becomes flatter, especially when t >= 50. The signal modulator exp(-t/200) is more straight within this range. Within the range of 1 <= t <= 100, the first signal modulator shows more exponential appearance, and it decreases faster compared to the second one.

## (d)
For series in (a) and (b), the mean functions can be computed as follow:

E(x~t~) = E(s~t~ + w~t~) = E(s~t~) + E(w~t~) = s~t~ + 0 = s~t~, since w~t~ is white noise with mean 0.

Thus the mean function of the two series are just the s~t~ defined in (a) and (b) respectively.

The plots of mean functions are as follow:
```{r, fig.height=3.8}
# plot the mean functions of (a)
plot.ts(s_a, main = "Mean of Question 1(a) Series", col = "blue")
# plot the mean functions of (b)
plot.ts(s_b, main = "Mean of Question 1(b) Series", col = "blue")
```

\newpage

# Question 2
## (a)
The regression model is fitted as follow:
```{r}
#install.packages("astsa")
library(astsa)
```

```{r}
trend <- time(jj) - 1970 # center the time and make the regression coefficients look nicer
Qtr <- factor(cycle(jj)) # make Qtr factor with 4 levels
reg_model <- lm(log(jj) ~ 0 + trend + Qtr, na.action = NULL) 
# the 0 ensures that no intercept for this model
summary(reg_model) # summarize the regression model
```
The estimated model is: x~t~ = 0.167172t + 1.052793Q~1~(t) + 1.080916Q~2~(t) + 1.151024Q~3~(t) + 0.882266Q~4~(t)

All the five estimated parameters are significant since they all have p-value 
< 2e-16, we have stong evidence against the null hypothesis that the true parameter is 0. The full model is also significant since it also has p-value < 2e-16.

\newpage
## (b)
Since we have Gaussian white noise, the true mean annual increase in the logged earmings per share is: 

$E(x_{t+1}-x_t) = E(\beta (t+1) + w_{t+1} - \beta t -w_t) = \beta$

And according to the output, the estimated average annual increase is 0.167172 in the logged earmings per share.

## (c)
In the 3rd season: Q~3~(t) = 1 and others are 0

$E(x_t) = E(\beta t + \alpha_3 + w_t) = \beta t + \alpha_3$

In the 4th season: Q~4~(t) = 1 and others are 0

$E(x_t) = E(\beta t + \alpha_4 + w_t) = \beta t + \alpha_4$

Thus the change from the 3rd quarter to the 4th quarter is:

$\beta t + \alpha_4 - \beta t - \alpha_3 = \alpha_4 - \alpha_3$ = 0.882266 - 1.151024 = -0.268758

The average logged earning rate decreased by 0.268758 from the 3rd quarter to the 4th quarter. 
And it decreased by 0.268758/1.151024 = 23.35%


\newpage
## (d)
```{r}
reg_int <- lm(log(jj) ~ trend + Qtr) # include an intercept for this model
summary(reg_int) # summarize the regression model
```

After including an intercept term in the model, the model no longer have the coefficient for the first quarter, and the estimated coefficient for the second quarter becomes not significant because of the large p-value. The estimated coefficient for the third quarter has a larger p-value too. These consequences are not desirable.

## (e)
```{r, fig.height=4}
# graph the Johnson&Johnson data
plot(log(jj), main = "The Actual Data vs. Fitted Values", 
     ylab = "Logged earnings per share")
# superimpose the fitted values in red
lines(fitted(reg_model), col = "red")
```
The actual data is in black, and the fitted value is in red.

```{r, fig.height=4}
# plot the residuals
plot.ts(resid(reg_model), main = "Residuals of the regression model", 
        ylab = "Residuals")
```

By observing the residual plot, we can see that there is pattern in it, thus the residuals do not look like white noise. Therefore the fitted regression model does not satisfy the assumption of white noise and does not fit this data well.

\newpage

# Question 3
## Determine the autocovariance and autocorrelation functions as a function of lag h = s - t (hand-writen).

\newpage
## The plot of the ACF is as follow:
```{r}
h <- c(-6:6) # the range of the lag
# the value of the ACF
acf_h <- c(0, 0, 0, 0, 1/3.44, 2.4/3.44, 1, 2.4/3.44, 1/3.44, 0, 0, 0, 0)
# the plot of ACF as a function of h
plot(h, acf_h, type = 'h', xlab = "h", ylab = "ACF", 
     main = "The ACF as a function of h")
```
















